A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution \thanks{A preliminary version of this work has appeared in the Proceedings of the 26th International Symposium on Graph Drawing and Network Visualization (GD 2018). This work is supported by the DFG grant Ka812/17-1}

### 2018-09-25 paper/journal.tex
new figs

% ==================================================================
\input{cr-res-rome.tex}
% ==================================================================

% ==================================================================
\input{to-res-rome.tex}
% ==================================================================

% ==================================================================
\input{an-res-rome.tex}
% ==================================================================

% ==================================================================
\input{cr-res-north.tex}
% ==================================================================

% ==================================================================
\input{to-res-north.tex}
% ==================================================================

% ==================================================================
\input{an-res-north.tex}
% ==================================================================

@@

\begin{table*}[t!]
\caption{Summary of the results for the Graph Drawing Contest 2017.}
\centering
\begin{tabular}{c|c|c|c|c}
\toprule
~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
\midrule
1 & 90  & 77  & 89.99 & 89.99\\
2 & 88.23  & 42 & 88.21 & 88.7 \\
3 & 90  & 89 & 87.86 & 89.95 \\
4 & 88.97  & 89 & 77.13  & 89.05 \\
5 & 80.4  &  30 & 78.68 & 86.96 \\
6 & 90  & 78 & 89.96 & 89.96\\
7 & 56.537  & 34 & 55.77  & 63.62 \\
8 & 84.95  & 61 & 81.18 & 89.28 \\
9 & 59.885  & 9 & 54.63 & 88.2 \\
10 & 20.978  & 4 & 23.60  & 23.72 \\
11 & 46.684  & 6 & 57.00 & 72.00 \\
12 & 36.47  & 5 & 26.24 & 35.86 \\
13 & 25.456  & 4 & 22.43 & 33.68 \\
14 & 33.52  & 5 & 29.69 & 43.08 \\
15 & 20.512  & 4 & 13.51 & 29.18 \\
\bottomrule
\end{tabular}
\label{tab:gdContest2017}
\end{table*}

@@

\begin{figure}[t]
\centering
\subfloat[\label{fig:graph5-yfiles}{}]{
\includegraphics[width=0.3\textwidth]{graph5-yfiles}}

\subfloat[\label{fig:graph9-yfiles}{}]{
\includegraphics[width=0.27\textwidth]{graph9-yfiles}}
\caption{Illustration of 
(a)~the 5th, and of
(b)~the 9th graph of the Graph Drawing contest 2017.}
\label{fig:graph5And9-yfiles}
\end{figure}

@@

%
In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing. Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio. 
%As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.

@@

We conclude by noting that our motivation to work with this problem was our participation to the Graph Drawing contest 2017, where we performed miserably using a force-directed algorithm. As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely better as also evidenced by our experiments. The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.

>>>>

% ==================================================================
%\input{cr-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=1]{figures/rome}}
	\hfil
	\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=2]{figures/rome}}

	\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=3]{figures/rome}}
	\hfil
	\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=4]{figures/rome}}
	\caption{Experimental results on the crossing resolution for the Rome graphs.}
	\label{fig:cr-res}
\end{figure*}

% ==================================================================
%\input{to-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:to-res-1}{Crossing resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=5]{figures/rome}}
	\hfil
	\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=6]{figures/rome}}

	\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=7]{figures/rome}}
	\hfil
	\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=8]{figures/rome}}
	\caption{Experimental results on the total resolution for the Rome graphs.}
	\label{fig:to-res}
\end{figure*}

% ==================================================================
%\input{an-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:an-res-1}{Crossing resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=9]{figures/rome}}
	\hfil
	\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=10]{figures/rome}}

	\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=11]{figures/rome}}
	\hfil
	\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=12]{figures/rome}}
	\caption{Experimental results on the angular resolution for the Rome graphs.}
	\label{fig:an-res}
\end{figure*}

% ==================================================================
%\input{cr-res-north.tex}
% ==================================================================
\begin{figure*}[t!]
	\centering
	\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=1]{figures/north}}
	\hfil
	\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=2]{figures/north}}

	\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=3]{figures/north}}
	\hfil
	\subfloat[\label{fig:northCrossing-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=4]{figures/north}}
	\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
	\label{fig:northCrossing}
\end{figure*}

% ==================================================================
%\input{to-res-north.tex}
% ==================================================================
\begin{figure*}[t!]
	\centering
	\subfloat[\label{fig:northTotal-1}{Crossing resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=5]{figures/north}}
	\hfil
	\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=6]{figures/north}}

	\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=7]{figures/north}}
	\hfil
	\subfloat[\label{fig:northTotal-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=8]{figures/north}}
	\caption{Experimental results for the total resolution on the AT\&T graphs.}
	\label{fig:northTotal}
\end{figure*}

% ==================================================================
%\input{an-res-north.tex}
% ==================================================================
\begin{figure*}[t!]
	\centering
	\subfloat[\label{fig:northAngular-1}{Crossing resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=9]{figures/north}}
	\hfil
	\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=10]{figures/north}}

	\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=11]{figures/north}}
	\hfil
	\subfloat[\label{fig:northAngular-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=12]{figures/north}}
	\caption{Experimental results for the angular resolution on the AT\&T graphs.}
	\label{fig:northAngular}
\end{figure*}

@@

\begin{table*}
\caption{Summary of the results for the Graph Drawing Contest 2017.}
\centering
\begin{tabular}{c|c|c|c|c}
\toprule
~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
\midrule
1 & 90.00  & 77  & 89.99 & 89.99\\
2 & 88.23  & 42  & 88.21 & 88.7 \\
3 & 90.00  & 89  & 87.86 & 89.95 \\
4 & 88.97  & 89  & 77.13 & 89.05 \\
5 & 80.40  & 30  & 78.68 & 86.96 \\
6 & 90.00  & 78  & 89.96 & 89.96\\
7 & 56.53  & 34  & 55.77 & 63.62 \\
8 & 84.95  & 61  & 81.18 & 89.28 \\
9 & 59.88  & 9   & 54.63 & 88.2 \\
10 & 20.97  & 4  & 23.60 & 23.72 \\
11 & 46.68  & 6  & 57.00 & 72.00 \\
12 & 36.47  & 5  & 26.24 & 35.86 \\
13 & 25.45  & 4  & 22.43 & 33.68 \\
14 & 33.52  & 5  & 29.69 & 43.08 \\
15 & 20.51  & 4  & 13.51 & 29.18 \\
16 & 19.83  & 3  & 12.13 & 26.12 \\
\bottomrule
\end{tabular}
\label{tab:gdContest2017}
\end{table*}

@@

\begin{figure}[t]
\centering
\subfloat[\label{fig:graph5-yfiles}{}]{
\includegraphics[width=0.4\textwidth]{graph5-yfiles}}

\subfloat[\label{fig:graph9-yfiles}{}]{
\includegraphics[width=0.37\textwidth]{graph9-yfiles}}
\caption{Illustration of 
(a)~the 5th, and of
(b)~the 9th graph of the Graph Drawing contest 2017.}
\label{fig:graph5And9-yfiles}
\end{figure}

@@

%
In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing. Our experimental evaluation indicates that the new heuristic is competitive to the state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.

@@

We conclude by noting that our motivation to work with this problem was our participation to the Graph Drawing contest 2017, where we performed miserably using a force-directed algorithm. As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely better as also evidenced by our experiments. The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing. Thus, as a future direction, we plan to further extend and evaluate our technique, e.g., by using a multi-criteria objective function (i.e., that is not limited to the crossing angles) that takes into account more aesthetic criteria.
